---
title: "Machine Learning, 2019/2020 - Assignment 2"
author: "Martin Geletka, Alessandra Crippa, Giulio Rago"
date: "June 03, 2020"
output: html_document
---


## Introduction
The purpose of this assignment is  to use generated artificial data to
study some properties of a few ANN architectures. The studied architectures are

* MLPs
* RNNs
* LSTMs

In this assignment, we will compare the behaviour of these mehtods in terms of accuracy and convergence and show the best results for each network


## Generating pseudo XML data
```{r}
gen_tag <- function(maxtaglen=5, all_letters=1) {
  taglen<-floor(runif(1)*maxtaglen+1)
  paste(letters[sample(1:(length(letters)*all_letters),taglen,replace=TRUE)], collapse='')
}

gen_line <- function(depth=5,...) {
  inner <- ""
  size <- runif(1)*7+3
  tag<-gen_tag(...)
  inner <-""
  while(runif(1)>0.4 && depth>0)
    if(inner=="")
      inner<-gen_line(depth=depth-1,...)
    else
      inner<-paste(inner,gen_line(depth=depth-1,...))
      endtag<-paste("<",tag,collapse='',sep='')
    if(inner=="")
      paste(tag,endtag)
    else
      paste(tag,inner,endtag)
}
# this function generates full sequences. You can truncate them using indexes.
gen_xml_data <- function(len=1000,...)
  sapply(1:len,function(i) gen_line(...)
         )

data <- gen_xml_data(len=20, depth=2,maxtaglen=3,all_letters=0.3)
data
```
###MLP - TODO
```{r}

```

###RNN - TODO
```{r}

```

###LSTM -TODO
```{r}

```

